---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# trackballr
<!-- badges: start -->
[![DOI](https://zenodo.org/badge/773406370.svg)](https://zenodo.org/doi/10.5281/zenodo.13235277)
[![R-CMD-check](https://github.com/roaldarbol/trackballr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/roaldarbol/trackballr/actions/workflows/R-CMD-check.yaml)
[![trackballr status
badge](https://roaldarbol.r-universe.dev/badges/trackballr)](https://roaldarbol.r-universe.dev)
[![Codecov test coverage](https://codecov.io/gh/roaldarbol/trackballr/graph/badge.svg)](https://app.codecov.io/gh/roaldarbol/trackballr)
<!-- badges: end -->

*Run the trackball!*

Trackball experiments, in which animals are tethered/restrained atop a (most commonly) styrofoam ball are common experiments within animal behaviour and neuroscience.

The primary goal of the *trackballr* package is to enable a *tidyverse*-friendly analysis of trackball data and use a standardised  tidy data format.

## Installation

You can install the development version of trackballr with:

Type | Source | Command
---|---|---
Development | R-universe | `install.packages("trackballr", repos = "https://roaldarbol.r-universe.dev")`

Once you have installed the package, you can load it with:

```{r, eval=FALSE}
library("trackballr")
```

## Documentation

The trackball analysis workflow is simple, and generally follows the same steps as other movement data sets (e.g. pose estimation, centroid tracking). See our docs to go through the steps, one-by-one:

- [Read data](https://www.roald-arboel.com/trackballr/articles/Read-data.html)
- [Clean tracks](https://www.roald-arboel.com/trackballr/articles/Clean-tracks.html)
- [Compute kinematics](https://www.roald-arboel.com/trackballr/articles/Compute-kinematics.html)
- [Compute movement statistics](https://www.roald-arboel.com/trackballr/articles/Compute-movement-statistics.html)

## Status
> [!Warning]
>
> 🏗️ The package is currently in early development and the interface is subject to change. Feel free to play around and provide feedback.


## Roadmap

We envision this package to develop mostly into providing functions for reading trackball data. We hope to support your experimental cases, such as:

- ✅ Support for Optical mouse sensors, free configuration
- ⏳ Support for Optical mouse sensors, fixed configuration
- ⏳ Support FicTrac
- ⏳ More tests

If you are running experiments with trackballs, we would love to get a sample of your data to support it!

For computing kinematics and summary statistics, we are currently implementing those, but they might be transferred to a different package that does so for a broad array of movement data, so **stay put to hear more**!

## Citation

To cite *trackballr* in publications use:

```{r}
citation("trackballr")
```
