---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# trackballr
<!-- badges: start -->
[![DOI](https://zenodo.org/badge/773406370.svg)](https://zenodo.org/doi/10.5281/zenodo.13235277)
[![R-CMD-check](https://github.com/roaldarbol/trackballr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/roaldarbol/trackballr/actions/workflows/R-CMD-check.yaml)
[![trackballr status
badge](https://roaldarbol.r-universe.dev/badges/trackballr)](https://roaldarbol.r-universe.dev)
[![Codecov test coverage](https://codecov.io/gh/roaldarbol/trackballr/graph/badge.svg)](https://app.codecov.io/gh/roaldarbol/trackballr)
<!-- badges: end -->

*Run the trackball!*

Trackball experiments, in which animals are tethered/restrained atop a (most commonly) styrofoam ball are ...
The primary goal of the *trackballr* package is to make analysis of trackball data easy and standardise the data format.  

## Installation

You can install the development version of trackballr with:

Type | Source | Command
---|---|---
Development | R-universe | `install.packages("trackballr", repos = "https://roaldarbol.r-universe.dev")`

Once you have installed the package, you can load it with:

```{r, eval=FALSE}
library("trackballr")
```

## Documentation

For worked examples of how to use the package, see the vignettes:

- [Analyse single trial](https://www.roald-arboel.com/trackballr/articles/Analyse-Single-Trial.html)


## Citation

To cite *trackballr* in publications use:

```{r}
citation("trackballr")
```
